---
title: 'Exposome: supervised analysis of the transcriptomics data'
author: "Julien Chiquet"
date: "June 27, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, message = FALSE, warning = FALSE)
```


## Required packages

```{r}
library(tidyverse)
library(glmnet)
library(stabs)
```

## Data preparation

Import the data
```{r}
covariates  <- read_delim("Covariates.txt" , delim = "\t") %>% dplyr::select(-subjectidp1)
exposures   <- read_delim("Exposures.txt"  , delim = "\t") %>% dplyr::select(-subjectidp1)
transcripts <- readRDS("Transcripts.rds") %>% as_tibble(rownames = "subjectidp")
```

Merge covariates and exposures and retain rows shared by all tables
```{r}
covar_expos <- left_join(covariates, exposures, by = "subjectidp")
transcripts <- semi_join(transcripts, covar_expos, by = "subjectidp")
```

## Principal component Analysis

A basic PCA indicates that the city variable has a strong effect:
```{r}
library(FactoMineR)
pca_transcrit   <- PCA(dplyr::select(transcripts, -subjectidp), graph = FALSE)
```

```{r, echo = FALSE}
colors <- factor(covar_expos$city)

par(mfrow=c(2,2))
barplot(pca_transcrit$eig[1:10, 3], las = 3)
plot(pca_transcrit, axes = c(1,2), choix = "ind", col.ind = colors)
plot(pca_transcrit, axes = c(1,3), choix = "ind", col.ind = colors)
plot(pca_transcrit, axes = c(2,3), choix = "ind", col.ind = colors)
```

## Filtering before predicting `modeledpm25`

I choose to predict the `modeledpm25` response. I thus filter the row corresponding to NA values in `modeledpm25` in both the covariates/exposome variables and the table of transcripts.

```{r}
covar_expo_transcript <- semi_join(
  filter(covar_expos, !is.na(modeledpm25)), transcripts
  )
expr <- dplyr::select(semi_join(transcripts, covar_expo_transcript, "subjectidp"), -subjectidp)
response <- pull(covar_expo_transcript, modeledpm25) 
```

## Variable screening: retain the most varying transcrit

Basic screening based on the variance of the transcrits, only keeping the first - says - 2000.
```{r}
nvar <- 2000
vars <- order(apply(as.matrix(expr), 2, var, na.rm = TRUE), decreasing = TRUE)
expr_sub <- as.matrix(expr[, vars[1:nvar]])
```

The design matrix is composed by the city variable (dummy coding) + the transcript.
```{r}
X <- cbind(model.matrix(~ covar_expo_transcript$city + 0), scale(expr_sub))
```

## Variable section with the LASSO

I try the LASSO without penalizing the city effect

```{r}
penalties <- c(rep(0, 4), rep(1, ncol(X)-4))
## now launch glmnet
out_cv_glmnet <- cv.glmnet(X, response, standardize = FALSE, penalty.factor = penalties)
```

```{r, echo = FALSE}
par(mfrow= c(1,2))
plot(out_cv_glmnet)
plot(out_cv_glmnet$glmnet.fit)
```

Obviously the transcripts effects are essentially masked by the city effect: we prefer performin variable seleciton on each data subset separatly.

## Variable selection per city

```{r}
expr_city <- lapply(split(expr_sub, covar_expo_transcript$city), matrix, ncol = ncol(expr_sub))
response_city <- split(response, covar_expo_transcript$city)
```

```{r}
out_cv_glmnet_city <- Map(cv.glmnet, expr_city, response_city)
```

```{r}
par(mfrow = c(2,4))
lapply(out_cv_glmnet_city, plot)
lapply(out_cv_glmnet_city, function(x) plot(x$glmnet.fit))
```

There seems to be some signal, so let us try some stability selection approach.

## Stability Selection

```{r}
my_stabs <- function(x, y) {
  stabsel(x = x, y = y, fitfun = glmnet.lasso, cutoff = 0.75, PFER = 1)
}
stability_sel <- Map(my_stabs, expr_city, response_city)
```

An the winners are
```{r}
print(lapply(stability_sel, function(x) colnames(expr_sub)[x$selected]))
```


